---
title: "_Data-Analysis-Replication_"
format: html
editor: visual
---

# Data Analysis Replication

## Classification pf psychedelics and psychoactive drugs based on brain-wide imaging of cellular. c-Fos expression

### Loading file to re-create table in figure 1-C

```{r}
f<-"https://raw.githubusercontent.com/meganhall719/data-analysis-replication/refs/heads/main/Data/HTR_summary_data.csv"
H<-read_csv(f,col_names = TRUE)
head(H)
summary(H)
```

##### In the data set total_HTR is head-twitch count we are going to pull out the individual data points too make the data easy to work with

##### To keep data organized I am using H in front of all my drug conditions for all my head count data

```{r}
library(ggplot2)

# Define the drugs vector - note that H_DET has 6 entries, not 7
drugs <- c(rep("H_Sal", 7), rep("H_DET", 6), rep("H_PSI", 6), rep("H_5_MEO", 6)) 
           
head_twitch_counts <- c(
  # Saline data
  14, 43, 20, 16, 26, 18, 18,
  # 6-F-DET data
  18, 47, 27, 27, 29, 58,
  # Psilocybin data
  73, 63, 61, 51, 75, 85,
  # 5-MEO data
  89, 85, 68, 52, 87, 45)

# Verify lengths
length(drugs)  # Should be 25
length(head_twitch_counts)  # Should be 25

# Create the dataframe
data <- data.frame(
  Drug = drugs,
  head_twitch_count = head_twitch_counts
)

# Adding colors
drug_colors <- c(
  "H_Sal" = "grey",
  "H_DET" = "lightblue",
  "H_PSI" = "green",
  "H_5_MEO" = "blue"
)

head(data)

#making boxplots
data$Drug <- factor(data$Drug, levels = c("H_Sal", "H_DET", "H_PSI", "H_5_MEO"))

p <- ggplot(data, aes(x = Drug, y = head_twitch_count, fill = Drug)) +
  geom_boxplot(outlier.shape = NA, width = 0.7) +
  geom_point(position = position_jitter(width = 0.2, seed = 123), 
             color = "black", size = 2.5) +
  geom_point(position = position_jitter(width = 0.2, seed = 123), 
             color = drug_colors[as.character(data$Drug)], size = 2) +
  
  scale_fill_manual(values = drug_colors) +
  scale_y_continuous(
    name = "Head-twitch count",
    limits = c(0, 150),
    breaks = seq(0, 150, by = 50)
  ) +
  theme_classic() +
  theme(
    axis.title.y = element_text(face = "bold", size = 12),
    axis.text = element_text(color = "black", size = 10),
    axis.title.x = element_blank(),
    axis.text.x = element_text(face = "bold", size = 11, color = c("grey", "lightblue", "green", "blue")),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "none"
  )

# Print the plot
print(p)
```

### Loading file to re-create table in figure 1-D

```{r}
# load in file 
library(tidyverse)

f<-"https://raw.githubusercontent.com/meganhall719/data-analysis-replication/refs/heads/main/Data/data%20analysis%20repo%20c-Fos_CellCounts_Compiled.csv"
d<- read_csv(f, col_names = TRUE)
head(d)
```

#### Organize data from data frame so it is workable

Taking sum of each drug conditions total c-Fos+ cells and taking their average

-   There are 8 drug conditions each drug condition was repeated 8 times for a total of n= 64

-   The number of cFos+ Cells detected in each brain region was listed in the data frame, the sum aka total number of cFos+ cells were pulled out from the data frame for each trial i.e (drug condition x 1-8) and the average was taken for each drug condtion

```{r}
PSI <-c(1273778, 2326276,1557597,3267826, 2070772, 2618733, 1727876, 2008088)
avg_PSI <- mean(PSI, na.rm = TRUE)


KET <-c(2813047, 2996137, 2750500,6174431,2802245,3985839,2804512,5141505)
avg_KET <-mean(KET, na.rm = TRUE)

fiveMEO<- c(2874169,2741911,6084926,2172441,2213387,3073534,3023897,2393271)
avg_5MEO <-mean (fiveMEO, na.rm = TRUE)

DET <- c(1540624,2028787,1904801,2710060,5213341,2439566,2969437,4878214)
avg_DET <-mean(DET, na.rm = TRUE)

MDMA <- c(4962783,4802318,3654292,4707371,4150196,4712711,4743384,5019285)
avg_MDMA <- mean (MDMA, na.rm =  TRUE)

aSSRI<- c(1532878,2040636,2414929,2538250,2479113,2394598,1650509,1562409 )

avg_aSSRI <-mean(aSSRI, na.rm = TRUE)

C_SSRI <- c(878021,1326870,1151260,1254341,1167137,1462502,1555244,1031251)
avg_C_SSRI<- mean(C_SSRI, na.rm = TRUE)

SAL<-c(3433408,2408396,2498773,916741,2420846,1841269,1139141,3856055)
avg_SAL <-mean(SAL, na.rm = TRUE)

avg_PSI
avg_KET
avg_5MEO
avg_DET
avg_MDMA
avg_aSSRI
avg_C_SSRI
avg_SAL
```

#### Creating a visual for cFos+ positive cell counts across all drug conditions

```{r}
# Load required libraries
library(ggplot2)
library(tidyr)
library(dplyr)

# Create a dataframe with your data
# organizing data in a tidy format
drugs <- c(rep("PSI", 8), rep("KET", 8), rep("5MEO", 8), rep("DET", 8), 
           rep("MDMA", 8), rep("A-SSRI", 8), rep("C-SSRI", 8), rep("SAL", 8))

# Combineining all cell count data 
cell_counts <- c(
  # PSI data
  1273778, 2326276, 1557597, 3267826, 2070772, 2618733, 1727876, 2008088,
  # KET data
  2813047, 2996137, 2750500, 6174431, 2802245, 3985839, 2804512, 5141505,
  # 5MEO data
  2874169, 2741911, 6084926, 2172441, 2213387, 3073534, 3023897, 2393271,
  # DET data
  1540624, 2028787, 1904801, 2710060, 5213341, 2439566, 2969437, 4878214,
  # MDMA data
  4962783, 4802318, 3654292, 4707371, 4150196, 4712711, 4743384, 5019285,
  # A-SSRI data
  1532878, 2040636, 2414929, 2538250, 2479113, 2394598, 1650509, 1562409,
  # C-SSRI data
  878021, 1326870, 1151260, 1254341, 1167137, 1462502, 1555244, 1031251,
  # SAL data
  3433408, 2408396, 2498773, 916741, 2420846, 1841269, 1139141, 3856055
)

#Adding gender to data points
gender <- c(
  # PSI gender 
  "Male", "Male", "Male", "Male", "Female", "Female", "Female", "Female",
  # KET gender 
   "Male", "Male", "Male", "Male", "Female", "Female", "Female", "Female",
  # 5MEO gender
   "Male", "Male", "Male", "Male", "Female", "Female", "Female", "Female",
  # DET gender
   "Male", "Male", "Male", "Male", "Female", "Female", "Female", "Female",
  # MDMA gender
   "Male", "Male", "Male", "Male", "Female", "Female", "Female", "Female",
  # A-SSRI gender
   "Male", "Male", "Male", "Male", "Female", "Female", "Female", "Female",
  # C-SSRI gender
   "Male", "Male", "Male", "Male", "Female", "Female", "Female", "Female",
  # SAL gender
   "Male", "Male", "Male", "Male", "Female", "Female", "Female", "Female"
)

# Create the dataframe
data <- data.frame(
  Drug = drugs,
  Count = cell_counts,
  Gender = gender
)


# Adding colors
drug_colors <- c(
  "PSI" = "green",     # Green
  "KET" = "purple",     # Purple/pink
  "5MEO" = "blue",    # Blue
  "DET" = "lightblue",     # Light blue
  "MDMA" = "gold",    # Yellow/gold
  "A-SSRI" = "red",  # Red/orange
  "C-SSRI" = "lightpink",  # Light pink
  "SAL" = "gray"      # Gray
)

# Order the drugs as they appear in figure 1D
data$Drug <- factor(data$Drug, levels = c("PSI", "KET", "5MEO", "DET", "MDMA", "A-SSRI", "C-SSRI", "SAL"))

# Create the plot
p <- ggplot(data, aes(x = Drug, y = Count, fill = Drug)) +
  # Add the boxplot
  geom_boxplot(outlier.shape = NA, width = 0.7) +  
  # Add individual points with gender differentiation
  geom_point(aes(shape = Gender), position = position_jitter(width = 0.15), size = 2, alpha = 0.7) +
  # Add error bars (whiskers)
  stat_boxplot(geom = "errorbar", width = 0.4) +  
  # Use log scale for y-axis 
  scale_y_log10(
    name = "c-Fos+ cell count",
    breaks = c(1e6, 2e6, 5e6, 1e7),  
    labels = expression(10^6, 2 %*% 10^6, 5 %*% 10^6, 10^7),  # Formatted labels
    limits = c(8e5, 1e7)
  ) +
  # Apply drug-specific colors
  scale_fill_manual(values = drug_colors) +
  # Apply different shapes for gender
  scale_shape_manual(values = c("Female" = 4, "Male" = 16)) +  # X for female, dot for male
 
  theme_bw() +  
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5, face = "bold"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.y = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.position = c(0.95, 0.9),
    legend.background = element_rect(fill = "white", color = "black")
  ) +
  
  guides(fill = "none")


print(p)

```
