---
title: "Maternal Seperation"
format: html
editor: visual
---


##  Figure 1J

### Investigating how social defeat stress affects mitochondrial function in medial pre-frontal cortex MPFC

Accelerated social defeat stress(AcSD) paradigm they put their wild-type mice through it s a model used during discrete time periods 35-44 days old

experimenters then identified susceptible mice (Sus), resilient mice (Res), and Control mice (Con) that received normal rearing practices

```{r}

library(tidyverse)
f<-"https://raw.githubusercontent.com/meganhall719/_Data-Analysis-Replication_/refs/heads/main/Data/Fig.1j_mPFC.csv"

j<-read_csv(f,col_names = TRUE )

head (j)
```

```{r}
# Assuming your data is already loaded as 'j'
# First, rename the first column if needed
j <- j %>% rename(Subject = `...1`)

# Then transform to long format
j_long <- j %>%
  pivot_longer(
    cols = c(Con, Sus, Res),
    names_to = "Group",
    values_to = "Value"
  )

# Convert Group to a factor with meaningful labels
j_long <- j_long %>%
  mutate(Group = factor(Group, 
                      levels = c("Con", "Sus", "Res"),
                      labels = c("Control", "Susceptible", "Resilient")))

# View the transformed data
head(j_long)

# Calculate summary statistics
group_stats <- j_long %>%
  group_by(Group) %>%
  summarise(
    n = n(),
    mean = mean(Value, na.rm = TRUE),
    sd = sd(Value, na.rm = TRUE),
    se = sd / sqrt(n),
    ci_lower = mean - 1.96 * se,
    ci_upper = mean + 1.96 * se
  )

# View summary statistics
print(group_stats)

names(j_long)
str(j_long)
head(j_long)
```

```{r}
library(tidyverse)



# Calculate summary statistics for each group
group_stats <- j_long %>%
  group_by(Group) %>%
  summarise(
    mean = mean(Value, na.rm = TRUE),
    se = sd(Value, na.rm = TRUE) / sqrt(n()),
    n = n()
  )

# Create the bar graph
ggplot() +
  # Add bar graph showing means
  geom_bar(data = group_stats, 
           aes(x = Group, y = mean, fill = Group),
           stat = "identity", width = 0.6, alpha = 0.8) +
  
  # Add individual data points
  geom_point(data = j_long, 
             aes(x = Group, y = Value, color = Group),
             position = position_jitter(width = 0.2, seed = 123),
             size = 2, alpha = 0.7) +
  
  # Add error bars (standard error)
  geom_errorbar(data = group_stats,
                aes(x = Group, ymin = mean - se, ymax = mean + se),
                width = 0.2, linewidth = 0.8) +
  
  # Set colors
  scale_fill_manual(values = c("Control" = "gray", 
                               "Susceptible" = "pink", 
                               "Resilient" = "lightblue")) +
  scale_color_manual(values = c("Control" = "black", 
                                "Susceptible" = "red", 
                                "Resilient" = "blue")) +
  
  # Add group sizes at the bottom
  annotate("text", x = 1:3, y = 5, 
           label = paste0("n=", group_stats$n),
           size = 3.5) +
  
  # Customize labels and theme
  labs(y = "Mitochondrial ATP nmol/ug protien, (%)", x = "") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    plot.margin = margin(30, 10, 30, 10)
  )

```


## Running statistical analysis and following the papers methods one-way ANOVA and Bonferroni’s multiple comparison test

### These results matched the graph in the paper lets add our signifcant bars now

p = 0.0003 for Control vs. Susceptible p = 0.0017 for Susceptible vs. Resilient

```{r}
library(tidyverse)

# Using the exact data from the image
j <- data.frame(
  Subject = 1:6,
  Con = c(76.30973, 124.37320, 116.40440, 109.08300, 72.67641, 101.15330),
  Sus = c(44.74942, 65.50968, 58.40964, 42.68800, 36.61544, 26.20278),
  Res = c(92.26998, 58.22349, 98.37781, 96.88507, 89.84742, 113.31700)
)

# Transform to long format for analysis
j_long <- j %>%
  pivot_longer(
    cols = c(Con, Sus, Res),
    names_to = "Group",
    values_to = "Value"  # Note: using "Value" as the column name
  ) %>%
  mutate(Group = factor(Group, levels = c("Con", "Sus", "Res")))

#check the structure of j_long to confirm column names
str(j_long)

# Print summary statistics for each group
group_stats <- j_long %>%
  group_by(Group) %>%
  summarise(
    n = n(),
    mean = mean(Value),
    sd = sd(Value),
    se = sd / sqrt(n)
  )
print("Group statistics:")
print(group_stats)

# Run one-way ANOVA
anova_result <- aov(Value ~ Group, data = j_long)
anova_summary <- summary(anova_result)
print("ANOVA result:")
print(anova_summary)

# Run pairwise comparisons with Bonferroni correction
pairwise_result <- pairwise.t.test(
  j_long$Value, 
  j_long$Group,
  p.adjust.method = "bonferroni"
)
print("Pairwise comparisons with Bonferroni correction:")
print(pairwise_result)

# Try one-tailed t-tests (if the researchers had directional hypotheses)
# Control vs Susceptible (expecting Control > Susceptible)
t_con_sus <- t.test(
  j_long$Value[j_long$Group == "Con"],
  j_long$Value[j_long$Group == "Sus"],
  alternative = "greater"  # One-tailed test
)
print("One-tailed t-test, Control > Susceptible:")
print(t_con_sus$p.value)

# Susceptible vs Resilient (expecting Susceptible < Resilient)
t_sus_res <- t.test(
  j_long$Value[j_long$Group == "Sus"],
  j_long$Value[j_long$Group == "Res"],
  alternative = "less"  # One-tailed test
)
print("One-tailed t-test, Susceptible < Resilient:")
print(t_sus_res$p.value)

# Control vs Resilient
t_con_res <- t.test(
  j_long$Value[j_long$Group == "Con"],
  j_long$Value[j_long$Group == "Res"],
  alternative = "two.sided"  # Two-tailed test
)
print("Two-tailed t-test, Control vs Resilient:")
print(t_con_res$p.value)
```
